### 2015-03-17 stat310/project.tex
write up

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    \E[\hat{\beta} \vert X] &= \E[(X'X)^{-1}X'Y \vert X] \\
                                            &= \E[(X'X)^{-1}X'(X\beta+e) \vert X] \\
                                            &= \E[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
                                            &= \E[\beta \vert X] + (X'X)^{-1}X' \E[e \vert X] \\
                                            &= \beta + (X'X)^{-1}X' b(X)
                \end{align*}
                Thus, $\hat{\beta}$ is only unbiased when \((X'X)^{-1}X' b(X) = 0\).
                In other words, $\hat{\beta}$ is biased if  $b(X)$ is in the span of \((X'X)^{-1}X'\).
            \item[b)]
                \begin{align*}
                    \var[\hat{\beta} \vert X ] &= \var[(X'X)^{-1}X'Y \vert X ] \\
                                               &= \var[(X'X)^{-1}X'(X\beta+e) \vert X ] \\
                                               &= \var[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
                                               &= \var[\beta+(X'X)^{-1}X'e \vert X] \\
                                               &= \var[(X'X)^{-1}X'e \vert X] \\
                                               &= (X'X)^{-1}X'\var[e \vert X]((X'X)^{-1}X')' \\
                                               &= (X'X)^{-1}X'\sigma^2I X (X'X)^{-1} \\
                                               &= (X'X)^{-1}X'X(X'X)^{-1} \sigma^2 \\
                                               &= (X'X)^{-1}\sigma^2 \\
                \end{align*}
                Thus, though $\hat{\beta}$ may be biased, it has no higher variance with assumption A2$'$ than with the usual assumption A2.
    \item[2.]
        \begin{itemize}
            \item[i.]
                Given that the variable \(wt_i = \beta_0 + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, and independent of the variable height, the probability that a model that included height would reduce the residual sum of squares as much as it has or more is $17.31\%$.
            \item[ii.]
                Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_0 \vert \geq 36.8759 \) given a null hypothesis that \(\beta_0=0\) is $58.3\%$.
           \item[iii.]
               Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_1 \vert \geq .5821  \) given a null hypothesis that \(\beta_1=0\) is $17.3\%$.
            \item[iv.]
                You can not compute the probability of the fitted model being the right one, only the probability of seeing the given data and estimates on it for various assumptions of the underlying process that generated it.
    \item[3.]
        All the individuals formulas for the desired quantities are: \\
        \[SSreg = \frac{R^2 RSS}{1 - R^2}= \frac{.2084 \times 351191.4}{1-.2084}=92456.14\]
        \[RSS=(RSE \times df)^2=(2.993 \times 198)^2=351191.4\]
        \[MSreg = \frac{SSreg}{regDF} = \frac{92456.14}{1}= 92456.14\]
        \[\hat\sigma^2 = \frac{RSS}{resDF} = \frac{351191.41}{198} = 1773.694\]
        \[F = \frac{MSreg}{\hat\sigma^2} = \frac{92456.14}{1773.694} = 1773.694 \]
        Which fit in the table as such: \\
        \begin{tabular}{l | l l l l l}
            Regression & df      & ss               & ms       & F     \\
            \hline
            Regression & regDF = 1       & SSreg = 92456.14 & MSreg = 92456.14          & F = 52.126 \\
            Residual   & resDf = 198     & RSS = 351191.4   & $\hat\sigma^2$ = 1773.694 &  
        \end{tabular}
    \item[4.]
        \begin{itemize}
            \item[a)] These are the resulting confidence intervals: \\
                \begin{tabular}{l | r l r l}
                    Confidence Level & $\beta_0$ LB & $\beta_0$ UB & $\beta_1$ LB & $\beta_1$ UP \\
                    \hline
                    66\%             &       (3.802,& 3.828)       &       (-.062,& -.059)     \\
                    90\%             &       (3.792,& 3.837)       &       (-.063,& -.058)     \\
                    95\%             &       (3.788,& 3.841)       &       (-.064,& -.057)
                \end{tabular}
            \item[b)] We get the variance estimate:
            \[1.538\times10^{-5} = \sigma^2 \left[ \begin{array}{c c} 1 & 10 \end{array} \right] \left(X'X\right)^{-1} \left[ \begin{array}{c} 1 \\  10 \end{array} \right] \]
            \item[c)] This is what we end up with, for a 95\% CI: \\
                \begin{tabular}{l | r c l}
                    Type       & LB    & Fit   & UB    \\
                    \hline
                    Mean       & 3.005 & 3.022 & 3.038 \\
                    Prediction & 2.800 & 3.022 & 3.245 
                \end{tabular}
            \item[d)]
                The confidence interval is smallest near the mean of fixed acidity ($8.31$ in the data set). This stands to reason, since the confidence intervals are based on conditioned slices of a multivariate normal distribution; these slices will be steepest towards the center of the distribution (around the means of all variables), and flattest in the tails. In other words, its harder to screw up a prediction where you have a lot of data than where you don't have much data. \\
                \begin{center}
                    \includegraphics[width=12cm]{hw2_plot} 
                \end{center}
        Here is my R code: \\
        \lstinputlisting{hw2.R}

>>>>

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 310 Programing Assignment \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        After randomly generating $A$ and $c$, setting \({\bf b}=\frac{1}{2}A{\bf 1}\), and solving the continuous relaxation of the boolean linear programing problem, I got a \({\bf y}^*\). It only had 16 entries with values that weren't 0 or 1. \({\bf c}^T{\bf y}=-35.396\), which we know is a lower bound for \({\bf c}^T{\bf x}\), the optimal objective value for the original problem. I generated ${\bf x}_\tau$ for $\tau = .01n, n=1$ to $100$, and plotted the objective value and least slack among the rows of $A{\bf x}_\tau\leq{\bf y}$ below. 
        \begin{center}
            \includegraphics[width=12cm]{plot_1.pdf}
        \end{center}
        The feasible points lie to the left of $0$, corresponding to $A{\bf x}_\tau-{\bf y}$ having all negative entries(and thus the constraint being satisfied). Since each point in this region is a feasible point for the original boolean program, the corresponding objective value is an upper bound for the minimum objective value, since the minimum must be less than or equal to each of them. The smallest of these values is ${\bf c}^T{\bf x}=-34.373$. Thus, we have an upper bound on the optimal point of $-34.373$ and a lower bound of $-35.396$, the difference being $1.023$. \\

{\bf Note:} Since there are only $16$ values in ${\bf y}$ which aren't either $0$ or $1$, there could be at most unique 17 ${\bf x}_\tau$ (there are actually only 16 since two of the points which aren't 0 or 1 are within .01 of each other). This is because ${\bf x}_\tau = {\bf x}_{\tau+.01}$ if none of the entries of ${\bf y}$ lie in $(\tau,\tau+.01)$. \\

        {\bf Code:} \\
        \lstinputlisting[firstline=17,lastline=61]{ajmaurer_project.m}

    \item[2.]
        If we set \(A={\bf aa}'\), this problem can be equivalently written as, where \(X\in\mathbb{S}^n\),
        \begin{align*}
            \text{minimize    }\; & tr(AX) \\
            \text{subject to  }\; & X \succeq 0 \\
                                  & X =\left[ \begin{array}{cccc} .2  & x_1 & x_2 & x_3 \\
                                                                  x_1 & .1  & x_4 & x_5 \\
                                                                  x_2 & x_4 & .3  & x_6 \\
                                                                  x_3 & x_5 & x_6 & .1 \end{array} \right]  \\
                                  & x_1 \geq 0 & x_2 \geq 0 \\
                                  & x_4 \leq 0 & x_5 \leq 0 \\
                                  & x_6 \geq 0
        \end{align*}
        Plugging this into our software, we get that
        \[X^* = \left[ \begin{array}{cccc} .2    & 0   & .2449 &   0 \\
                                            0    & .1  &    0  & -.1 \\
                                           .2449 & 0   & .3    &   0 \\
                                            0    & -.1 &    0  &   .1 \end{array} \right],
         \; tr(AX^*) = .0013 \]
        Its interesting to note that our minimal objective value, even with the restrictions placed on $X$, is quite close to the minimum value you would get for arbitrary $X\succeq0$ of 0.

        {\bf Code:} \\
        \lstinputlisting[firstline=66,lastline=84]{ajmaurer_project.m}
